Score-based tests of measurement invariance: Use in practice

### 2014-01-26 frontiers/frontiers.Rnw
work on the comments and Rnw file.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@306 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The family of tests described in this paper have recently been applied
to the study of measurement invariance \cite{MerZei13,MerFanZei}, though their
practical application has been limited to a small set of simulations and
datasets.  In this paper, we provide a detailed illustration of the tests' use
and performance under
scenarios likely to be encountered in practice. 
%In order to fix the psychometric scale, we need to specify the violating parameter.  
%Moreover, in practice model misspecification is an inevitable issue to a certain extent 
%and may interfere with parameter invariance specificity. A family of tests derived from 
%stochastic process has been recently proposed to identify specific parameters impacted 
%by measurement invariance violation. However, these tests have only been used to a limited 
%type of parameter violation. It is this paper's intent to provide a detailed examination 
%of the tests' performance under practical research scenarios. 
In the following sections, we first briefly review the  theoretical framework of
the proposed tests and provide a short tutorial illustrating the use of the
tests in \proglang{R} \cite{R11}. Next, we study the tests' performance in simulations that mimic
practical research scenarios.  Finally, we provide some further discussion on
the tests' use in practice.

@@

To conduct the analysis, we first load the data and keep only complete cases for simplicity 
(though the tests can be applied to incomplete data).
%
<<data>>=
data("YouthGratitude", package = "psychotools")
compcases <- apply(YouthGratitude[, 4:28], 1, function(x) all(x %in% 1:9))
yg <- YouthGratitude[compcases, ]
@
%
Next, we fit two models in \pkg{lavaan}: a one-factor model 
where loadings are restricted to be equal across age groups, 
and a one-factor model where loadings are free across age groups.
This allows us to test a hypothesis of partial measurement invariance
that was of interest to the original researchers (though, for ordinal
variables, all types of
measurement invariance can be examined via the tests described
previously).
By default, the code below sets the scale by fixing the first loading
to 1.
%
<<models>>=
restr <- cfa("f1 =~ gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
  data = yg, group = "agegroup", meanstructure = TRUE,
  group.equal = "loadings")

>>>>

The family of tests described in this paper have recently been applied
to the study of measurement invariance \cite{MerZei13,MerFanZei}, though their
practical application has been limited to a small set of simulations and
datasets.  In this paper, we provide a detailed illustration of the tests' use
and performance under
scenarios likely to be encountered in practice.  To our knowledge, this 
is the first detailed examination of the tests' properties. \seeme{add
this sentence to point the unique contribution.}
%In order to fix the psychometric scale, we need to specify the violating parameter.  
%Moreover, in practice model misspecification is an inevitable issue to a certain extent 
%and may interfere with parameter invariance specificity. A family of tests derived from 
%stochastic process has been recently proposed to identify specific parameters impacted 
%by measurement invariance violation. However, these tests have only been used to a limited 
%type of parameter violation. It is this paper's intent to provide a detailed examination 
%of the tests' performance under practical research scenarios. 
In the following sections, we first briefly review the  theoretical framework of
the proposed tests and provide a short tutorial illustrating the use of the
tests in \proglang{R} \cite{R11}. Next, we study the tests' performance in simulations that mimic
practical research scenarios.  Finally, we provide some further discussion on
the tests' use in practice.

@@

To conduct the analysis, we first load the data and keep only complete cases for simplicity 
(though the tests can be applied to incomplete data).
%
<<data>>=
data("YouthGratitude", package = "psychotools")
compcases <- apply(YouthGratitude[, 4:28], 1, function(x) all(x %in% 1:9))
yg <- YouthGratitude[compcases, ]
@
%
Next, we fit two models in \pkg{lavaan}: a one-factor model 
where loadings are restricted to be equal across age groups, 
and a one-factor model where loadings are free across age groups.
This allows us to test a hypothesis of partial\seeme{I
think partial should be weak} measurement invariance
that was of interest to the original researchers (though, for ordinal
variables, all types of
measurement invariance can be examined via the tests described
previously).
By default, the code below sets the scale by fixing the first loading
to 1.
%
<<models>>=
restr <- cfa("f1 =~ gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
  data = yg, group = "agegroup", meanstructure = TRUE,
  group.equal = "loadings")
